---
title: "Drug Synergy RNA-seq analysis circos plots"
author: "Bobby Shih"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

Generate circos and intersection plots for RNA-seq from T-cell lymphoma cell lines treated with DMSO, Drug A, Drug B, or the combination to visualize and summarize similarities between different treatment conditions. 

```{r LOAD PACKAGES, message = FALSE}
# Load packages
library(DESeq2)
library(dplyr)
library(stringr)
library(pheatmap)
library(RColorBrewer)
library(ggplot2)
library(ggpubr)
library(reshape2)
library(circlize)
library(ComplexUpset)
```

Data processing for circos plot

```{r DATA PROCESSING AND PREPARATION}
# Use the below functions if below variables are not already in environment
drug_A_vs_dmso <- read.csv("/home/bobby/RStudio/RNA-seq/DrugSynergism/drug_A_vs_dmso.csv", row.names = 1)
drug_combo_vs_dmso <- read.csv("/home/bobby/RStudio/RNA-seq/DrugSynergism/drug_combo_vs_dmso.csv", row.names = 1)
drug_B_vs_dmso <- read.csv("/home/bobby/RStudio/RNA-seq/DrugSynergism/drug_B_vs_dmso.csv", row.names = 1)

# Matrix of up- and down-regulated genes in each possible comparison to DMSO
dysregulated.genes <- vector(mode = "list")
dysregulated.genes[["A_Upregulated"]] <- subset(drug_A_vs_dmso, log2FoldChange >= 1.2 & padj < 0.05) %>% row.names()
dysregulated.genes[["B_Upregulated"]] <- subset(drug_B_vs_dmso, log2FoldChange >= 1.2 & padj < 0.05) %>% row.names()
dysregulated.genes[["Combo_Upregulated"]] <- subset(drug_combo_vs_dmso, log2FoldChange >= 1.2 & padj < 0.05) %>% row.names()
dysregulated.genes[["A_Downregulated"]] <- subset(drug_A_vs_dmso, log2FoldChange <= -1.2 & padj < 0.05) %>% row.names()
dysregulated.genes[["B_Downregulated"]] <- subset(drug_B_vs_dmso, log2FoldChange <= -1.2 & padj < 0.05) %>% row.names()
dysregulated.genes[["Combo_Downregulated"]] <- subset(drug_combo_vs_dmso, log2FoldChange <= -1.2 & padj < 0.05) %>% row.names()

# Generate matrix for dysregulated genes unique to one condition
unique.genes <- vector(mode = "list")
unique.genes.temp <- c(dysregulated.genes[["A_Upregulated"]], 
                       dysregulated.genes[["B_Upregulated"]], 
                       dysregulated.genes[["Combo_Upregulated"]]) %>% table() %>% as.data.frame() %>% subset(Freq == 1)
unique.genes[["upregulated"]] <- unique.genes.temp$. %>% as.character()

unique.genes.temp <- c(dysregulated.genes[["A_Downregulated"]], 
                       dysregulated.genes[["B_Downregulated"]], 
                       dysregulated.genes[["Combo_Downregulated"]]) %>% table() %>% as.data.frame() %>% subset(Freq == 1)
unique.genes[["downregulated"]] <- unique.genes.temp$. %>% as.character()
rm(unique.genes.temp)

# Generate matrix for upregulated circos plot
# Note that the numbers indicate the size of the node between two groups
# note that the color of the ribbons in the circos plot are determined by the row of the circos matrix

upregulated.circos.mat <- matrix(0, nrow = 3, ncol = 5)
rownames(upregulated.circos.mat) <- c("A", "B", "Combo")
colnames(upregulated.circos.mat) <- c("A + B", "B + Combo", "B + Combo", "Unique", "Common")

upregulated.common <- intersect(dysregulated.genes[["A_Upregulated"]], 
                                dysregulated.genes[["B_Upregulated"]]) %>% intersect(dysregulated.genes[["Combo_Upregulated"]])
upregulated.circos.mat[,"Common"] <- intersect(dysregulated.genes[["A_Upregulated"]], 
                                               dysregulated.genes[["B_Upregulated"]]) %>%
                                                  intersect(dysregulated.genes[["Combo_Upregulated"]]) %>% length()

temp <- intersect(dysregulated.genes[["A_Upregulated"]], 
                  dysregulated.genes[["B_Upregulated"]]) %in% upregulated.common
upregulated.circos.mat["A", "A + B"] <- sum(!temp)
temp <- intersect(dysregulated.genes[["A_Upregulated"]], 
                  dysregulated.genes[["Combo_Upregulated"]]) %in% upregulated.common
upregulated.circos.mat["A","B + Combo"] <- sum(!temp)
upregulated.circos.mat["A","Unique"] <- intersect(dysregulated.genes[["A_Upregulated"]], 
                                                  unique.genes[["upregulated"]]) %>% length()

upregulated.circos.mat["B","A + B"] <- upregulated.circos.mat["A","A + B"]
temp <- intersect(dysregulated.genes[["B_Upregulated"]], 
                  dysregulated.genes[["Combo_Upregulated"]]) %in% upregulated.common
upregulated.circos.mat["B","B + Combo"] <- sum(!temp)
upregulated.circos.mat["B", "Unique"] <- intersect(dysregulated.genes[["B_Upregulated"]], 
                                                   unique.genes[["upregulated"]]) %>% length()

upregulated.circos.mat["Combo","B + Combo"] <- upregulated.circos.mat["A","B + Combo"]
upregulated.circos.mat["Combo","B + Combo"] <- upregulated.circos.mat["B","B + Combo"]
upregulated.circos.mat["Combo","Unique"] <- intersect(dysregulated.genes[["Combo_Upregulated"]], 
                                                      unique.genes[["upregulated"]]) %>% length()

rm(temp)

# GeneCombote matrix for downregulated circos plot
downregulated.circos.mat <- matrix(0, nrow = 3, ncol = 5)
rownames(downregulated.circos.mat) <- c("A", "B", "Combo")
colnames(downregulated.circos.mat) <- c("A + B", "B + Combo", "B + Combo", "Unique", "Common")

downregulated.common <- intersect(dysregulated.genes[["A_Downregulated"]], 
                                  dysregulated.genes[["B_Downregulated"]]) %>% intersect(dysregulated.genes[["Combo_Downregulated"]])
downregulated.circos.mat[,"Common"] <- intersect(dysregulated.genes[["A_Downregulated"]], 
                                                 dysregulated.genes[["B_Downregulated"]]) %>% 
                                                    intersect(dysregulated.genes[["Combo_Downregulated"]]) %>% length()

temp <- intersect(dysregulated.genes[["A_Downregulated"]], 
                  dysregulated.genes[["B_Downregulated"]]) %in% downregulated.common
downregulated.circos.mat["A", "A + B"] <- sum(!temp)
temp <- intersect(dysregulated.genes[["A_Downregulated"]], 
                  dysregulated.genes[["Combo_Downregulated"]]) %in% downregulated.common
downregulated.circos.mat["A","B + Combo"] <- sum(!temp)
downregulated.circos.mat["A","Unique"] <- intersect(dysregulated.genes[["A_Downregulated"]], 
                                                    unique.genes[["downregulated"]]) %>% length()

downregulated.circos.mat["B","A + B"] <- downregulated.circos.mat["A","A + B"]
temp <- intersect(dysregulated.genes[["B_Downregulated"]], 
                  dysregulated.genes[["Combo_Downregulated"]]) %in% downregulated.common
downregulated.circos.mat["B","B + Combo"] <- sum(!temp)
downregulated.circos.mat["B", "Unique"] <- intersect(dysregulated.genes[["B_Downregulated"]], 
                                                     unique.genes[["downregulated"]]) %>% length()

downregulated.circos.mat["Combo","B + Combo"] <- downregulated.circos.mat["A","B + Combo"]
downregulated.circos.mat["Combo","B + Combo"] <- downregulated.circos.mat["B","B + Combo"]
downregulated.circos.mat["Combo","Unique"] <- intersect(dysregulated.genes[["Combo_Downregulated"]], 
                                                        unique.genes[["downregulated"]]) %>% length()

# Make necessary group variables for plotting
nm = unique(unlist(dimnames(upregulated.circos.mat)))
group = structure(c(rep("A", 3), rep("B", 5)), names = nm)

# grid.col is used to assign colors to different groups, If you want two groups to have the same color, assign it the same number in grid.col
# Each number corresponds to a different color, so changing the number and order of numbers can get different plots
grid.col = structure(2:9,
                     names = nm)
#
```

Data processing for intersection plot

```{r, message = FALSE}
# Upset plot for upregulated genes
all.up.genes <- c(dysregulated.genes[[1]], dysregulated.genes[[2]], dysregulated.genes[[3]])
all.up.genes <- all.up.genes[!duplicated(all.up.genes)]

upregulated.intersection <- matrix(nrow = length(all.up.genes), ncol = 3)
rownames(upregulated.intersection) <- all.up.genes
colnames(upregulated.intersection) <- c("A", "B", "Combo")

for (i in 1:3){
  for (j in 1:length(all.up.genes)){
    upregulated.intersection[j,i] <- sum(dysregulated.genes[[i]] == all.up.genes[j])
  }
}

# Turn 1 to TRUE and 0 to FALSE
upregulated.intersection <- upregulated.intersection == 1
upregulated.intersection <- as.data.frame(upregulated.intersection)

# Upset plot for downregulated genes
all.down.genes <- c(dysregulated.genes[[4]], dysregulated.genes[[5]], dysregulated.genes[[6]])
all.down.genes <- all.down.genes[!duplicated(all.down.genes)]
downregulated.intersection <- matrix(nrow = length(all.down.genes), ncol = 3)
rownames(downregulated.intersection) <- all.down.genes
colnames(downregulated.intersection) <- c("A", "B", "Combo")

for (i in 1:3){
  for (j in 1:length(all.down.genes)){
    downregulated.intersection[j,i] <- sum(dysregulated.genes[[i+3]] == all.down.genes[j])
  }
}

# Turn 1 to TRUE and 0 to FALSE
downregulated.intersection <- downregulated.intersection == 1
downregulated.intersection <- as.data.frame(downregulated.intersection)
#

```

FIG 3A : Significantly up-regulated gene circos plots + intersection plot

```{r FIG3A_TOP, fig.align = 'center', fig.height = 9}
par(cex = 1.5, mar = c(0, 0, 0, 0))
chordDiagram(upregulated.circos.mat, 
             group = group, 
             grid.col = grid.col, 
             transparency = 0.3, 
             annotationTrack = "grid",
             preAllocateTracks = list(track.height = max(strwidth(unlist(dimnames(upregulated.circos.mat))))))

# Adjust label direction according to width of sectors
circos.track(track.index = 1, panel.fun = function(x, y) {
    xlim = get.cell.meta.data("xlim")
    xplot = get.cell.meta.data("xplot")
    ylim = get.cell.meta.data("ylim")
    sector.name = get.cell.meta.data("sector.index")

    if(abs(xplot[2] - xplot[1]) < 10) {
        circos.text(mean(xlim), 
                    ylim[1], 
                    sector.name, 
                    facing = "clockwise",
            niceFacing = TRUE, 
            adj = c(0, 0.5), col = "black")
    } else {
        circos.text(mean(xlim), ylim[1], 
                    sector.name, 
                    facing = "inside", 
            niceFacing = TRUE, 
            adj = c(0.5, 0), col= "black")
    }
}, bg.border = NA)
circos.clear()
```
```{r FIG3A_BOT, fig.align = 'center', fig.width = 9}
upset(upregulated.intersection, colnames(upregulated.intersection),
      base_annotations = list('Intersection size' = intersection_size(counts = TRUE, text = aes(size = 8))),
        themes=upset_modify_themes(
          list(
            'intersections_matrix'= theme(text=element_text(size=25)),
            'overall_sizes'= theme(axis.text.x=element_text(angle=90, size = 15),
                                   text = element_text(size = 20)),
            "Intersection size" = theme(text = element_text(size = 20),
                                        axis.text.y = element_text(size = 20))
              )
    )
)

```

FIG 3B : Significantly down-regulated gene circos plots + intersection plot

```{r FIG3B_TOP, fig.align = 'center', fig.height = 9}
par(cex = 1.5, mar = c(0, 0, 0, 0))
chordDiagram(downregulated.circos.mat, 
             group = group, 
             grid.col = grid.col, 
             transparency = 0.3, 
             annotationTrack = "grid",
             preAllocateTracks = list(track.height = max(strwidth(unlist(dimnames(downregulated.circos.mat))))))

# Adjust label direction according to width of sectors
circos.track(track.index = 1, panel.fun = function(x, y) {
    xlim = get.cell.meta.data("xlim")
    xplot = get.cell.meta.data("xplot")
    ylim = get.cell.meta.data("ylim")
    sector.name = get.cell.meta.data("sector.index")

    if(abs(xplot[2] - xplot[1]) < 10) {
        circos.text(mean(xlim), ylim[1], sector.name, facing = "clockwise",
            niceFacing = TRUE, adj = c(0, 0.5), col = "black")
    } else {
        circos.text(mean(xlim), ylim[1], sector.name, facing = "inside", 
            niceFacing = TRUE, adj = c(0.5, 0), col= "black")
    }
}, bg.border = NA)
circos.clear()
```
```{r FIG3B_BOT, fig.align = 'center', fig.width = 9}
upset(downregulated.intersection, colnames(downregulated.intersection),
      base_annotations = list('Intersection size' = intersection_size(counts = TRUE, text = aes(size = 8))),
        themes=upset_modify_themes(
          list(
            'intersections_matrix'= theme(text=element_text(size=25)),
            'overall_sizes'= theme(axis.text.x=element_text(angle=90, size = 15),
                                   text = element_text(size = 20)),
            "Intersection size" = theme(text = element_text(size = 20),
                                        axis.text.y = element_text(size = 20))
              )
    )
)
```

FIG 3C : Heatmap of all significantly up- and down-regulated genes when comparing DMSO vs. Combination

```{r FIG3C Data, message = FALSE}

dds_drug <- readRDS("/home/bobby/RStudio/RNA-seq/DrugSynergism/DESEQ2_dds_drug.rds")

combo_vs_dmso <- results(dds_drug, name = "drug_combo_vs_dmso")

all_pairwise_siggenes <- combo_vs_dmso %>% subset(padj < 0.05 & abs(log2FoldChange) > 1.2) %>% rownames()

dds.norm.counts <- read.csv("/home/bobby/RStudio/RNA-seq/DrugSynergism/DESEQ2_NormalizedCounts.csv", row.names = 1)

dds.norm.counts <- dds.norm.counts[, c('DMSO.1', 'DMSO.2', 'DMSO.3', 
                                       'A.1', 'A.2', 'A.3', 
                                       'B.1', 'B.2', 'B.3', 
                                       'COMBO.1', 'COMBO.2', 'COMBO.3')]

colnames(dds.norm.counts) <- c("DMSO-1", "DMSO-2", "DMSO-3", 
                               "A-1", "A-2", "A-3", 
                               "B-1", "B-2", "B-3", 
                               "COMBO-1", "COMBO-2", "COMBO-3")

```

```{r FIG3C, fig.align = 'center', fig.width = 9, fig.height = 12}

annotation_col <- structure(list(category = structure(c(1L, 1L, 1L, 2L, 2L, 2L, 3L, 3L, 3L, 4L, 4L ,4L), 
                                                      .Label = c("DMSO", "A", "B", "Combo"), class = "factor")), 
                            class = "data.frame", row.names = c("DMSO-1","DMSO-2","DMSO-3", 
                                                                "A-1","A-2","A-3", 
                                                                "B-1","B-2","B-3", 
                                                                "COMBO-1","COMBO-2","COMBO-3"))

heatmap <- pheatmap(dds.norm.counts[match(all_pairwise_siggenes, row.names(dds.norm.counts)),],
         cluster_rows = TRUE,
         cluster_cols = FALSE,
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(100),
         show_rownames = FALSE,
         show_colnames = FALSE,
         treeheight_row = 75,
         fontsize_col = 15,
         fontsize_row = 6,
         annotation_col = annotation_col,
         main = "All Pairwise Sig Genes",
         angle_col = 45)


```

